
# QPM Lab 1: Installing R and Rstudio 

Department of Political Science 

Washington University in St. Louis 

(Date here)

## What is R? 

R can be used for many different purposes. It is a programming language, a letter, a piece of data analysis software, and a community with over 2 million users. R is open source and completely free, so if you ever encounter a problem, you can go online and the R community will try to help you out. 

![](/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/1.png)

R's open source nature also means that you can contribute! If you discover or develop something, the R community would love to learn about it. 

We will use R in our Lab sections, to complete problem sets, and you will use it to analyze data for the final presentation. Many large companies use R to analyze data, so this experience will be useful in the long run and we recommend maintaining your R knowledge even after this class is over. 

## Installing R 

1. Close any running applications on your computer.

2. To obtain R, go to The Comprehensive R Archive Network (CRAN) (http://www.r-project.org/)

3. Click "Download R".

4. Scroll down to the USA, choose "CRAN Mirror **United States,Washington University, St. Louis**".

5. You will see versions ofRfor Linux, MacOS, and Windows. Downloadand install the version that corresponds to your operating system.

6. Warning for Windows users: You might want to be careful with the directory you download to! 

This is what R should look like, once opened: 

![](/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/2.png)

## Installing Rstudio 

RStudio allows you, the user, to run R in a more user-friendly environment. Like R, it is open-source and free. To obtainRStudio, go to the following link: (http://www.rstudio.com/) 

1. Click on "Download RStudio"

2. Scroll down and choose your operating system (under the Installers block) 

This is what Rstudio should look like, once opened: 

![](/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/3.png) 

Here are some basic things to try in R: 

- Multiplying, dividing, adding numbers in Rstudio console: 

```{r, eval=FALSE}
5*3
5+3
5/3
```

- Assigning a value to an object: (This assigns x the value of 6) 

```{r, eval=FALSE}
x <- 6 
```

- Install a package:

```{r, eval=F}
install.packages("foreign")
```

- Call a package through its library: 
```{r, eval=F}
library(foreign)
```

This CRAN-package, foreign, will come handy when you read data into R next week!

Here are some additional resources:
 
An introduction to R: (http://cran.r-project.org/doc/manuals/R-intro.pdf) 

An introduction to RStudio: (http://dss.princeton.edu/training/RStudio101.pdf) 