# Lab 4: Who's Winning the Election? 

Department of Political Science 

Washington University in St. Louis 

(date here) 

##Class Goals 

- Distribution functions in R

- CI in R basics

- More on plotting normal distributions

- More advanced subsetting

- Use statistical knowledge to assess the 2016 election

##Distribution Functions 

- "d"ensity - given $x$, returns height of a density curve.

```{r}
dnorm(.6, mean = 0, sd = 1)
```

- "q"uantile - for given p, returns the $x$ such that p% of the curve is to the left of $x$.

```{r}
qnorm(c(0.025, 0.975), mean = 0 , sd = 1)
```

- p"robability -  for given $x$, returns the probability of drawing $x$ or smaller from the distribution.

```{r}
pnorm(-1.96, mean = 0, sd = 1)
```

- "r"andom - draw $n$ values from the distribution at random.

```{r}
rnorm(1, mean = 0, sd = 1)
```

- For distributions other than Normal, change the suffix (e.g. for Student's $t$ distribution, {binom} for Binomial distribution)

## CI Basics

Recall, the formula for a CI _for a mean given sufficiently large n_ is 
$$\bar{y} \pm (Z\times \sigma_{\bar{y}})$$

We can build this in \texttt{R} as follows:

- bar_y <- mean(var) 

- sigma_bar_y <- sd(var)/sqrt(n)

- z <- qnorm(c(upper, lower), mean = 0, sd = 1) 


Putting it all together: 

- bar\_y + z*sigma\_bar\_y

## Plotting Two Normal Distributions on One Graph 

```{r}
ruler <- seq(0, 100, 1)
plot(x = ruler, y = dnorm(ruler, mean = 50, sd = 15), xlab = "x",ylab = "Density", xlim = c(0, 100), ylim = c(0, .045),type = "l", main = "Two Normal Distributions")
lines(x = ruler, y = dnorm(ruler, mean = 40, sd = 10), col = "blue")
```

##Plotting a Distribution Over a Histogram 

```{r}
hist( rnorm(500, mean = 6, sd = 2), ylim = c(0, .23),main = "Normal Distribution", xlab = "x",col = "aquamarine", freq = F)
ruler <- seq(0, 14, .01)
lines( x = ruler, y = dnorm(ruler, mean = 6, sd = 2),col = "darkmagenta", lwd = 2 )
```

##Advanced Subsetting 

How can we select observations that meet multiple conditions?

- == translates to “Is equal to"

- >,<,>=,<= translate to “greater than", "less than", etc.

- & translates to “and"

- | translates to “or"

Examples:

```{r, eval=F} 
dataname$varname[dataname$varname2 == value]#If value is text, put it in quotes!
dataname$varname[dataname$varname2 == value& dataname$varname3 == value2]
dataname[dataname$varname > value,  ] #all columns
dataname[dataname$varname == value |dataname$varname2 > value2, ]
```

## Sampling Distribution of a Proportion 

- For random sampling with a {\bf large} sample size n, the sampling distribution of the proportion $p$ is approximately normal.

- The mean of the sampling distribution is $p$ and standard deviation is $\sqrt{\dfrac{p(1-p)}{n}}$

- Hence, we can write $p \sim N\left(p,\sqrt{\dfrac{p(1-p)}{n}}\right)$

- If p is very close to 0 or 1, we need a larger sample for this statement to be true.
