# Lab 5: CI's in R 

Department of Political Science 

Washington University in St. Louis 

(date here) 

## Class Goals 

- Plotting the t distribution, with different degrees of freedom

- Building confidence intervals (CIs) 

- Working on lab handout

## Example Graphs and R Code: 

```{r}
par(mfrow=c(1,3))
elif <- seq(-6.869, 6.869, length.out=100)
plot(elif, dt(elif, df=2), xlab="x", ylab="Density",main=paste("t Distribution:  Degrees of freedom=2"),type="l",col="green",ylim=c(0,0.5))
plot(elif, dt(elif,df=10), xlab="x", ylab="Density",main=paste("t Distribution:  Degrees of freedom=10"),type="l",col="red",ylim=c(0,0.5))
plot(elif, dt(elif,df=30), xlab="x", ylab="Density",main=paste("t Distribution:  Degrees of freedom=30"),type="l",col="blue",ylim=c(0,0.5))
```

## What is a Confidence Interval? 

- Interval estimate of a population parameter

- Takes into account the uncertainty

- Depends on the data and confidence level

- Confidence level: if confidence intervals are constructed across manyseparate data analyses, the proportion of such intervals that containthe true value of the parameter will match the confidence level

- What happens when we increase/decrease confidence level?

```{r, echo=F, out.width = "400px"}
knitr::include_graphics("/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/19.png")
```

Figure :Normal population: 30 samples (n=150),μ=0, 90% confidence interval) 

Out of the 30 confidence intervals, only 27 of them (90%) include μ!

##Confidence Intervals for the Mean of Populations 

- Obtain $\bar{y}=\frac{\sum{y_i}}{n}$

- Obtain $\sigma_{\bar{y}}=\frac{\sigma}{\sqrt{n}}$ or $\hat{\sigma}_{\bar{y}}=\frac{S}{\sqrt{n}}$

- Choose a confidence level
 
- Obtain confidence coefficient: $\frac{\text{confidence level}}{100}$

- Find the Z-score that corresponds to $\frac{(\text{1-confidence coefficient})}{2}$ [Normal Table or R]


Confidence Interval for your chosen confidence level: 

($\bar{y}-z*\sigma_{\bar{y}},\bar{y}+z*\sigma_{\bar{y}}$)

## In R: 

###Numerical Summaries: 

- Get the mean, standard deviation and number of observations of thevariable you are interested in

###Normal Quantiles

- Gives you the required z-score for your confidence coefficient 

- Be sure tochoose the “upper tail”option so that your R Commanderresult matches the Normal table

### Code 

Obtain the Z- score and save it in an object called z.score: 

```{r, eval=F} 
z.score = qnorm(c((1-confidence coefficient)/2),lower.tail=FALSE))
```

Get mean, sd, and n of the variable and save in appropriately named objects:

```{r, eval=F}
mean.variable = mean(nameofdataset$nameofvariable,na.rm=T)
sd.variable = sd(nameofdataset$nameofvariable,na.rm=T)
n = length(na.omit(nameofdataset$nameofvariable))
```

Compute lower bound: 

```{r, eval=F}
mean.variable - z.score*sd.variable/sqrt(n)
```

Compute upper bound:

```{r, eval=F}
mean.variable+ z.score*sd.variable/sqrt(n)
```

Calculating 99% Confidence Interval for mean proportion of Clinton votefrom polls data (from last week): 

```{r, eval=F}
z.score = qnorm(c(0.005), lower.tail=FALSE))
n = length(na.omit(polls$Clinton))
mean.clinton = mean(polls$Clinton,na.rm=T)
sd.clinton = sd(polls$Clinton,na.rm=T)
lower.bound = mean.clinton -z.score*sd.clinton/sqrt(n)
upper.bound = mean.clinton +z.score*sd.clinton/sqrt(n)
```

The 99% Confidence Interval is: (42.827,44.091)

