# Lab 8: T-Tests In R 

Department of Political Science 

Washington University in St. Louis 

(date here)

## Class Goals 

- Recoding factors

- Hypothesis testing for the difference of means in R

## Gerber, Green, & Larimer GOTV Data 

```{r, eval=F}
social <- read.csv("https://raw.githubusercontent.com/
     kosukeimai/qss/master/CAUSALITY/social.csv")
colnames(social)
```

This should return: 

[1] "sex" "yearofbirth" "primary2004" "messages" "primary2006" "hhsize"   

In 2006, one of three mailers was sent out as part of a study:

- Civic Duty (It's your duty to vote)

- Hawthorne (You're being studied)

- Neighbors (Your neighbors will know if you voted)

How do these mailers impact turnout?

```{r, echo=F, out.width = "400px"}
knitr::include_graphics("/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/21.png")
``` 

We want to reduce messages to ``Treatment" and ``Control" :

```{r, eval=F}
table(social$messages)
```

This returns: 

Civic Duty    Control  Hawthorne  Neighbors 
     38218     191243      38204      38201 
     
```{r, eval=F}
install.packages("car")
library(car)
social$messages2 <- recode(social$messages, 
     " 'Hawthorne'='Treatment'; 'Civic Duty'='Treatment';
      'Neighbors'='Treatment' ")  
```

Does turnout differ between treatment and control?

```{r, eval=F}
t.test(outcome ̃group, alternative = "two.sided",var.equal = FALSE, data =dataset)
```

## t.test() Output

```{r, eval=F}
t.test(primary2006 ~ messages2, alternative = "two.sided",
     var.equal = FALSE, data = social)
```

Welch Two-Sample T-Test 

data:  primary2006 by messages2

t = -23.869, df = 234580, p-value < 2.2e-16

alternative hypothesis: true difference

      in means is not equal to 0
      
95 percent confidence interval:   -0.04506412 -0.03822505
 
sample estimates:

mean in group Control mean in group Treatment:  0.2966383      0.3382829 

## Options with the t.test function

Adjust for our hypothesis, given $H_0$ is $\mu_a = \mu_b$:

- if $H_a:$ $\mu_a \neq \mu_b$, then alternative = "two.sided"

- if $H_a:$ $\mu_a > \mu_b$, then alternative = "greater"

- if $H_a:$ $\mu_a < \mu_b$, then alternative = "less"

Adjust for variance assumptions:

- If we assume unequal variance, then \texttt{var.eq = F}

- If we assume equal variance, then \texttt{var.eq = T} 

## Alternative Usage of the t.test Function 

We can also give t.test two vectors.  First, create objects for turnout for two specific mailers:

```{r, eval=F}
neighbors <- social[social$messages == "Neighbors", "primary2006"]
civic_duty <- social[social$messages == "Civic Duty","primary2006"]
```

t.test(neighbors, civic_duty)

Welch Two Sample t-test

data:  neighbors and civic_duty

t = 18.463, df = 76271, p-value < 2.2e-16

alternative hypothesis: true difference in means is not equal to 0

95 percent confidence interval: 0.05667888 0.07014226

sample estimates:

mean of x: 0.3779482

mean of y: 0.3145377 

## Mailers from the 2016 Campaign 

```{r, echo=F, out.width = "400px"}
knitr::include_graphics("/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/22.jpg")
```

## Negative Advertising 

```{r, echo=F, out.width = "400px"}
knitr::include_graphics("/Users/Ian/Dropbox (IsD)/QPM/bookdown-minimal-master/bookdown-minimal-master/_bookdown_files/Lab_Files/23.png")
```

## Negative Advertising Experiment 

Voters were contacted prior to the 2004 election and randomly given either a \textbf{positive} or \textbf{negative} message on an issue.

For example, on education:

- Republicans failed to fund their own ``Leave No Child Behind" law and stuck Minnesotans with the cost... We deserve better leadership.

- Democrats plan to provide more resources, smaller class sizes and more textbooks to public schools....We need to insure that every child receives a good education. 

The authors then gathered turnout data for contacted voters in order to ask: Did the negative message impact turnout? 