## Lab 8 Handout: Quantitative Political Methodology

1. Write down the names of all the members of your group in this space.  We will collect one sheet from eachgroup for grading.

2. We  want  to  test  the  effect  of  negative  advertising  on  turnout  using  the  data  from  a  field  experiment by Arceneaux and Nickerson (Arceneaux,  Kevin and David W. Nickerson.  2010.  "Comparing Negativeand Positive Campaign Messages:  Evidence From Two Field Experiments." American Politics Research 38.).  You can download the data at (https://pages.wustl.edu/montgomery/negative-ads-dataset#overlay-context=montgomery).  The key variable of interest is a message delivered to voters by phone the weekend before the 2004 presidential election.  This is recorded in the data as treatment (1 = negative frame, 2 = positive frame). The outcome variable of interest is voted02p (0 = did not vote, 1 = voted). 

- Download the data from the QPM website and read it into R (Hint:  use the function read.delim()). 

```{r, eval=F}
neg <- read.delim("https://pages.wustl.edu/montgomery/negative-ads-dataset#overlay-context=montgomery")
```

- Overwrite the variable treatment as a factor with categories negative and positive (Hint:  use thefunction factor() that we covered last week.)

```{r, eval=F}
neg$treatment <- factor(neg$treatment, levels = c(1,2), labels = c("negative", "positive"))
```

3. Set up null and alternative hypotheses to find out whether negative messages make a difference for voter turnout.  Once you have done this, get the TA to initial your hypotheses before proceeding.

Null: Negative messages make no difference for voter turnout 

Alternative: Negative messages cause (increased/decreased) voter turnout.  

4. Given your hypotheses, conduct a t-test.  Write down your code and the output obtained.  Do negative messages make a difference for voter turnout after all?  Explain with reference to information from your R output.  How confident are you?

```{r, eval=F}
t.test(voted02p ~ treatment, data = neg)
```

5. If you have time, test your hypotheses for the younger voting population (under 25 years old).  First, subset your dataset.  Then conduct a t-test.  Write down your code and the output.  Interpret this output.  Donegative ads make a difference for younger voter turnout?  Explain.

```{r, eval=F}
young_neg <- neg$voted02p[neg$age <= 25 & neg$treatment == "negative"]
young_pos <- neg$voted02p[neg$age <= 25 & neg$treatment == "positive"]

t.test(young_neg, young_pos)
```

6. Compare your results from the two t-tests.  Did they produce the same output?